---
title: "Create models"
author: "Qian Zhao"
date: "11/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

In this markdown I illustrate the data generation procedure in Section 6.

The original data used to construct the HMM can be loaded as following: you can also download it from `SNPknock` package version 0.7.1 at `extdata/genotypes.RData`.

```{r}
load("./genotypes.RData")
table(X)
```

Convert X into the suitable fastPHASE input format, write it into a temporary file and return the path to that file.
```{r}
Xinp_file <- writeXtoInp(X)
```

Run fast phase on this synthetic dataset

```{r, eval}
fp_path  = "./fastPHASE_1" # Path to the fastPHASE executable
fp_outPath = runFastPhase(fp_path, Xinp_file, K = 12, numit = 15)

r_file = paste(fp_outPath, "_rhat.txt", sep="")
alpha_file = paste(fp_outPath, "_alphahat.txt", sep="")
theta_file = paste(fp_outPath, "_thetahat.txt", sep="")
char_file  = paste(fp_outPath, "_origchars", sep="")

hmm_1 <- loadHMM(r_file, alpha_file, theta_file, char_file, phased=FALSE, compact = TRUE)
hmm <-  loadHMM(r_file, alpha_file, theta_file, char_file, phased=TRUE, compact = FALSE)

# save R object 
# save(hmm_1, hmm, file = "./hmm.RData")
```

Alternatively, you can load the file that has been written as follows

```{r}
load("./hmm.RData")
```

Sample from the HMM 

```{r}
X = sampleHMM(hmm$pInit, hmm$Q, hmm$pEmit, n=5000)
dim(X)
```



