---
title: "Section 7"
author: "Qian Zhao"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align="center",fig.width=4, fig.height = 3)
```

Load required library, functions, and set up themes to produce plots. 

```{r, include = FALSE}
library(tidyverse)
library(lpSolve)
```

```{r, include= FALSE}
# global theme for plots
my_theme <- theme(
  axis.text.x = element_text(color = "grey10", size = 12),
      axis.text.y = element_text(color = "grey10", size = 12),
      axis.title.x = element_text(color = "black", size = 15),
      axis.title.y = element_text(color = "black", size = 15),
      plot.margin = unit(c(0.3, 0.3, 0.2, 0.2), "cm"))
```

```{r, include = FALSE}
# Function to detect whether data is separable
is.separable <- function(X, Y){
  # the code is adapted from https://rpubs.com/abhaypadda/linear-optimization-example
  X <- cbind(1, X)
  n <- dim(X)[1]; p <- dim(X)[2]
  # the objective is sum_i a_i x_i (x_i are the variables), then this is the vector of a_i
  objective.in <- colSums( X * matrix(rep(2*Y-1, each = p), n, p, byrow = TRUE))
  objective.in <- c(objective.in, -objective.in)
  # rhs for the constraints
  const.rhs <- c(rep(0, n), rep(1, 2*p))
  # create constraint martix
  const.mat <- rbind(X * matrix(rep(1-2*Y, each = p), n, p, byrow = TRUE),
                     diag(rep(1, p)),
                     diag(rep(-1, p)))
  const.mat <- cbind(const.mat, -const.mat)
  # constraints direction
  const.dir  <- rep("<=", (n + 2*p))
  # find the optimal solution
  optimum <- lp(direction="max",  objective.in, const.mat, const.dir,  const.rhs)
  z1 <- optimum$solution[1:p]
  z2 <- optimum$solution[(p+1):(2*p)]
  x <- z1 - z2
  
  ifelse(sum(x^2)<10^(-5), FALSE, TRUE)
}
```

### Load data

The UCI wine quality data contains 4898 observations and 11 covariates, including fixed acidity, volatile acidity, citric acid etc. 

```{r}
wine.url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv"
wine <- read.csv(wine.url, header = TRUE, sep = ";")
```

### Preprocessing

```{r}
wine$quality <- ifelse(wine$quality<=5, 0, 1) # "high" quality wine has score higher than 5
wine[,c(4, 6)] <- log(wine[,c(4, 6)]) # log-transform to make covariates look more normally distributed
wine[,-12] <- scale(wine[,-12], center = TRUE, scale = FALSE) # centering
```

```{r}
p <- dim(wine)[2] - 1 # number of covariates 
n <- dim(wine)[1] # number of observations
```

### MLE on the full data

```{r}
fit_full <- glm(quality~., data = wine, family = binomial(link = "logit"))
beta <- fit_full$coef
```
### MLE on sub-samples 

Focus on the variable - volatile acidity

```{r}
v <- 2
```

```{r}
kappas <- c(0.02, 0.06, 0.10, 0.14, 0.18, 0.22, 0.26) # problem dimensions
ns <- round(p/kappas)
```

```{r}
set.seed(2020101) # make sure we get the same answer
```

```{r}
B <- 100 # repeat B times 

mle <- matrix(0, B, length(kappas)) # fitted MLE
sd_r <- matrix(0, B, length(kappas)) # standard error from R
sd_fisher <- numeric(length(kappas)) # inverse Fisher info, plug in beta_hat from full data

for(i in 1:length(kappas)){
  nn <- round(p/kappas[i])
  b <- 1
  while(b<=B){ # retain values when the data is not separable
     data <- wine[sample(1:n, nn, replace = TRUE), ]
     
     emp <- matrix(0, p+1, p+1) 
     if(!is.separable(as.matrix(data[,-12]), data[,12])){
         fit <- glm(quality~., data = data, family = binomial(link = "logit"))
         mle[b,i] <- coef(fit)[v+1] # fitted coefficient
         sd_r[b,i] <- summary(fit)$coef[v+1,2] * sqrt(nn); # standard error from R, adjusted by sample size
         # standard error from Fisher information
         X <- as.matrix(cbind(1,data[,-12]));  
         eta <- as.vector(X%*%beta)
         W <- diag(1/(1+exp(eta))/(1+exp(-eta)))
         
         emp <- emp + t(X)%*%(W%*%X)
         b <- b+1
     }
  }
     avg=emp/B
     sd_fisher[i]=sqrt(solve(avg)[v+1,v+1])
}
```

### Median of the MLE 

```{r}
apply(mle, 2, median)
apply(mle, 2, median)/beta[3]
```

### Plots - fitted coefficients

Boxplot of the fitted coefficients

```{r}
boxdata <- tibble(
  kappa = factor(rep(kappas, each = B)),
  mle = as.vector(mle),
  sd_r = as.vector(sd_r) 
)
```

```{r}
g_coef <- boxdata %>% 
  filter(mle<20 & mle > -35) %>% 
  ggplot(.) +
  geom_boxplot(aes(x = kappa, y=mle)) + 
  geom_hline(yintercept = fit_full$coefficients[v+1],linetype="longdash", color = "red", size = 0.6) + 
  scale_y_continuous(breaks = seq(-30, 20, by=10), limits = c(-42, 25))+
  xlab(expression(kappa))+
  ylab("Estimated Coefficient")+
  theme_bw()+
  my_theme

# add the points that has been thresholded
text1 <- boxdata %>% 
  filter(mle>=20) %>% 
  group_by(kappa) %>% 
  summarize(n = n(),
            val = round(max(mle))) %>% 
  mutate(n = map_chr(n, ~paste0("[", ., "]")))

text2 <- boxdata %>% 
  filter(mle<=-35) %>% 
  group_by(kappa) %>% 
  summarize(n = n(),
            val = round(min(mle))) %>% 
  mutate(n = map_chr(n, ~paste0("[", ., "]")))

# add range
g_coef <- g_coef + 
  geom_text(aes(x = text1$kappa, y = 20,  label = text1$n),data = text1, size = 4) +
  geom_text(aes(x = text1$kappa, y = 25,  label = text1$val),data = text1, size = 4) +
  geom_text(aes(x = kappa, y = -37,  label = n), data = text2, size = 4) +
  geom_text(aes(x = kappa, y = -42,  label = val), data = text2, size = 4) 

g_coef
```

Saving figure
```{r}
filename <- "/real/figure/g_coef.png"
ggsave(filename, plot = g_coef, scale = 1, width = 5, height = 3, units = "in", dpi = 300) 
```


### Standard deviation from partitioning

```{r}
# n - total number of observations
# size - size of each partition
partition <- function(n, size){ 
  nn <- floor(n / size)
  matrix(sample(1:n, n, replace=FALSE)[1:(nn*size)], nn, size)
}
```

```{r}
sd_true <- numeric(length(kappas))

for(i in 1:length(kappas)){
  nn <- round(p/kappas[i])
  sample_index <- partition(n, nn)
  
  mle <- NULL
  for(b in 1:dim(sample_index)[1]){
    data <- wine[sample_index[b,], ]
    if(!is.separable(as.matrix(data[,-12]), data[,12])){
        fit <- glm(quality~., data = data, family = binomial(link = "logit"))
        mle <- c(mle, coef(fit)[v+1])
     }
  }
  sd_true[i] <- sd(mle) * sqrt(nn)
}
```

### Plots - standard error

```{r}
g_sd_r <- ggplot()+
  geom_boxplot(aes(x = kappa, y=sd_r), data = boxdata %>% filter(sd_r<120)  ) + 
  geom_point(aes(x=factor(kappas),y = sd_true), color = "red", shape = 17, size = 2)+
  xlab(expression(kappa))+
  ylab("Estimated standard error")+
  scale_y_continuous( breaks = c(25, 50, 75, 100, 125, 150), limits = c(10, 150))+
  theme_bw()+
  my_theme
  
# add the points that has been thresholded
text1 <- boxdata %>% 
  filter(sd_r >= 120) %>% 
  group_by(kappa) %>% 
  summarize(n = n(),
            val = round(max(sd_r))) %>% 
  mutate(n = map_chr(n, ~paste0("[", ., "]")))

# add range
g_sd_r <- g_sd_r + 
  geom_text(aes(x = text1$kappa, y = 140,  label = text1$n),data = text1, size = 4) +
  geom_text(aes(x = text1$kappa, y = 150,  label = text1$val),data = text1, size = 4) 

g_sd_r
```

```{r}
g_sd_fisher <- ggplot()+
    geom_point(aes(x = factor(kappas), y=sd_fisher),color = "blue", shape = 17, size = 2) + 
    geom_point(aes(x=factor(kappas),y =sd_true), color = "red", shape = 17, size = 2)+
   scale_y_continuous( breaks = c(25, 50, 75, 100, 125, 150), limits = c(10, 150))+
    xlab(expression(kappa))+
    ylab("Estimated standard error")+
    theme_bw()+
    my_theme
  

g_sd_fisher
```

Save Figure

```{r}
filename <- "/real/fig/g_sd_r.png"
ggsave(filename, plot = g_sd_r, device = "png", scale = 1, width = 4, height = 3, units = "in", dpi = 300) 

filename <- "/real/fig/g_sd_fisher.png"
ggsave(filename, plot = g_sd_fisher, device = "png", scale = 1, width = 4, height = 3, units = "in", dpi = 300) 
```
