---
title: "Logistic MLE"
author: "Qian Zhao"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
par(mar=c(3,3,2,1), mgp=c(1.5,0.5,0))
```

```{r, echo = FALSE}
logistic <- function(t) 1/(1+exp(-t))
```

We simulate a small dataset containing 400 observations $(X_i,Y_i)$, where each row of the covariates $X_i\in\mathbb{R}^{80}$ are independently sampled from a multivariate Gaussian $X_i\sim\mathcal{N}(0, \Sigma)$. $\Sigma$ is that of an $AR(1)$ model with correlation $\rho = 0.8$,i.e.
\[
\Sigma_{ij} = \rho^{|i-j|}.
\]
The response variable $Y_i$ is related to $X_i$ via a logistic model 
\[
Y_i\,|\, x_i\sim \mathrm{Bernoulli}(\mathrm{logistic}(X_i^\top \beta)),
\]
$\mathrm{logistic}(t) = 1/(1+e^{-t})$.  Here, we choose coefficients $\beta_i$ such that half of them are zero and the other half are equal with magnitude 0.2 (signal strength parameter $\gamma = 1.42$). 

```{r}
n <- 400
p <- 80
kappa <- p/n

Sigma <- toeplitz(0.2^(0:(p-1))) # covariance matrix
R <- chol(Sigma)

non_null <- sample(1:p, p/2, replace = FALSE) 
beta <- numeric(p)
beta[non_null] <- 0.2
```

### Logistic MLE

Let's look at the distribution of a single MLE $\hat{\beta}_{j}$ by repeated sampling. 

```{r}
B <- 1000
betas <- matrix(0, B, p)

for(i in 1:B){
  X <- matrix(rnorm(n*p, 0, 1), n, p)  %*% R
  Y <- rbinom(n, 1, logistic(X%*%beta))
  fit <-  glm(Y~X+0, family = binomial)
  betas[i,] <- coef(fit)
}
```

Plot a histogram of the fitted in all these repetitions. The estimated MLE is not centered at the true $\beta_j$ (red line), instead, it is biased upward in magnitude. 

```{r}
j <- sample(non_null,1)
hist(betas[,j], xlab = "Fitted MLE", main = "")
abline(v = beta[j], lty = "dashed", col = "red", lwd = 2)
abline(v=1.36*beta[j], lty = "dashed", col = "blue", lwd = 2)
```

The blue line plots $\alpha_\star\beta_j$, $\alpha_\star = 1.36$ is the predicted bias from the asymptotic theory. It is roughly at the center of all the fitted MLEs.

### Confidence intervals 

According to the asymptotic theory, the confidence intervals of the form 
\[
\left[\frac{1}{\alpha_\star}\left(\hat{\beta}_j - \frac{\sigma_\star}{\sqrt{n}\tau_j}z_{(1-\alpha/2)}\right),\frac{1}{\alpha_\star}\left(\hat{\beta}_j + \frac{\sigma_\star}{\sqrt{n}\tau_j}z_{(1-\alpha/2)}\right)\right]
\]
where $\tau_j = \mathrm{Var}(X_{ij}\,|\,X_{i,-j})$, should cover the true coefficient $\beta_j$ about $(1-\alpha)$ of the times. The pair of parameters $(\alpha_\star, \sigma_\star, \lambda_\star)$ solves a system of non-linear equations detailed in the paper (see also Sur (2019)). 

```{r}
# calculated parameters in this example
alpha_s <- 1.36
sigma_s <- 3.64
lambda_s <- 1.98
```

For this example, we can compute the conditional variance $\tau_j$ explicitely because the covariance matrix is known.

```{r}
tau <- 1/sqrt(diag(solve(Sigma)))
```

As an illustration, we compute a 90\% confidence interval each time, thus altogether there are $B$ confidence intervals.  

```{r}
# adjusted 90% confidence interval 
ci.low <- (betas + sigma_s / sqrt(n) / matrix(rep(tau, n), B, p, byrow = TRUE) * qnorm(0.05))/alpha_s
ci.up <-  (betas + sigma_s / sqrt(n) / matrix(rep(tau, n), B, p, byrow = TRUE)  * qnorm(0.95))/alpha_s
```

First, let's look at how often the $j$th variable is covered. 

```{r}
mean((beta[j]> ci.low[,j]) & (beta[j]< ci.up[,j])) 
```

Then, we compute the average coverage proportion in each repetition and report the average in $B$ repetitions. Both are quite close to 90\%. 

```{r}
mean((matrix(rep(beta, n), B, p, byrow = TRUE) > ci.low) & (matrix(rep(beta, n), B, p, byrow = TRUE) < ci.up)) 
```

### Further Reference
Pragya Sur and Emmanuel J. Candès, "A modern maximum-likelihood theory for high-dimensional logistic regression", *Proceedings of the National Academy of Sciences* Jul 2019, 116 (29) 14516-14525; DOI: 10.1073/pnas.1810420116


